---
title: "Exponential Smoothing forecast"
output: 
  html_notebook:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "hide"
    theme: "cosmo"
    highlight: "monochrome"
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#install.packages("devtools")
library(devtools)
library(tidyverse)
library(xtable)
library(knitr)
#library(ggpubr)
library(DT)
library(ggrepel)
library(forecast)

# Change the presentation of decimal numbers to 4 and avoid scientific notation
library(readxl)
options(prompt="R> ", digits=7, scipen=999)
```


# loading crime data

```{r}
library(readr)
long_master_nona <- read_csv("01-raw-data/noNA_long_all_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"),    col_types = cols(NDVr = col_double(), 
       eb_ndvr = col_double(), eb_npir = col_double(), 
       nimr = col_double(), Litr= col_double(), 
        nmrpr = col_double(), nmpr = col_double() ))

wide_master_nona <- read_csv("01-raw-data/noNA_wide_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"))
```

#creating and ID 1:1120 for each municipality time series

```{r}
nmr<- long_master_nona %>% 
  filter(year>=2003) %>% 
  select(2:3,year, starts_with("eb_nmr")) %>% 
  arrange(code) %>% 
  mutate(code=as.character(code))
#nmr %>% 
a<- nmr %>% 
  filter(year==2010) %>% 
  select(code) %>% 
  mutate(ID=1:nrow(.))
a
nmr<- left_join(nmr,a,by="code")
nmr
#nmr %>% group_by(ID) %>% summarise(n=n_distinct(municipality))
```

# creaing a list of time series

```{r}
ts_list=list()
id=c(1:1120)

for (x in seq_along(id)) {
  try <- nmr %>% 
  filter(ID==id[x])
  
  ts_list[[x]] <- ts(try$eb_nmr, start=2003)
}

```

# plotting the forecast for 2022 for one municipality, the correlogram and the box.Test


```{r}

id=171

ts_list[[id]]
autoplot(ts_list[[id]])

  ets(ts_list[[id]]) %>% 
forecast(h=4) %>% 
    autoplot()
  
ggAcf(ts_list[[id]], lag.max = 15)

Box.test(ts_list[[id]], lag=5, type = "Ljung")


```

    
# Estimating an ETS model for each municipality    
    
    
```{r}
nmr2022=c()
fit_ets=list()
for (x in 1:1120) {
  fit <- ets(ts_list[[x]])
#summary(fit)
 foreca <-fit %>% forecast(h=4)
nmr2022[x]= data.frame(foreca)[4,1]
fit_ets[[x]]<- fit
 }
# foreca is a dataframe with the forecast from 2019-2022  

method_ets=c()
for (x in 1:1120) {
 method_ets[x]<- fit_ets[[x]]$method
}


```
    
# type of ETS fit models 

```{r}
data_frame(id=1:1120,method_ets)
```


Residuals

Hyndman, R. J., & Athanasopoulos, G. (2018). Forecasting: principles and practice. OTexts.
3.3 book: "We suggest using h=10 for non-seasonal data and h=2m for seasonal data, where m is the period of seasonality  However, the test is not good when h is large, so if these values are larger than T/5, then use h=T/5"


```{r}
id=1

pval=c()

for (x in 1:1120) {
  f<-Box.test(residuals(fit_ets[[x]]),lag=3, type = "Lj")
pval[x]<- f$p.value
}
sum(pval<=0.05)

#pval<= 0.05
id=171
ggAcf(residuals(fit_ets[[id]]), lag.max = 15)
Box.test(residuals(fit_ets[[id]]), lag=3 , type = "Lj")

checkresiduals(fit_ets[[id]])

```

    
    
# According to this rough prediction how many municipalities will reach the 2022 target?
    
```{r}
nmr2018<- nmr %>% filter(year==2018)

nmr2018$pred2022 <- nmr2022
nmr2018

target<- 10000-(23.23/10)
target


nmr2018 %>% 
  filter(eb_nmr<target)

nmr2018 %>% 
  filter(pred2022<target)


```

# Are the time series white noise?

```{r}
pval=c()

for (x in 1:1120) {
  f<-Box.test(ts_list[[x]],lag=3, type = "Ljung")
pval[x]<- f$p.value
}
sum(pval<=0.05)
#pval
#pval<=0.05
```


#time  series cross validation using Exponential smoothing state space models

```{r}
fets <- function(x, h) {
  forecast(ets(x), h = h)
}
```


```{r}
#tsCV(ts_list[[1]], naive, h=1)
e_ets=list()
for (jj in 1:1120) {
  e<- tsCV(ts_list[[jj]], initial = 4, fets,h=4)
e_ets[[jj]]<- e[,4]
}
e_ets[[1]]
```



```{r}
mean_ets=list()
for (i in 1:1120) {
  mean_ets[[i]]=abs(e_ets[[i]])
}
```


# new table


```{r}
new_ets=list()
for (i in 1:1120) {
  new_ets[[i]]=abs(e_ets[[i]])
}

for (xx in 1:1119) {
  new_ets[[xx+1]]= new_ets[[xx]]+new_ets[[xx+1]]
}
new_ets[[1120]]/1120
mean_etsv<- (new_ets[[1120]]/1120)[5:12]
mean_etsv
```

```{r}
new_ets=list()
for (i in 1:1120) {
  new_ets[[i]]=(e_ets[[i]])^2
}
for (xx in 1:1119) {
  new_ets[[xx+1]]= new_ets[[xx]]+new_ets[[xx+1]]
}
new_ets[[1120]]/1120
rmse_ets<- (new_ets[[1120]]/1120)[5:12]
rmse_ets<- sqrt(rmse_ets)
rmse_ets


cv_yearly<- as.data.frame( rbind(mean_etsv, rmse_ets))

methodv=c("ETS (MAE)","ETS (RMSE)")
names=seq(2011,2018,1)
names= as.character(names)
colnames(cv_yearly)<-names
cv_yearly<- cv_yearly %>% 
  mutate(method=methodv) %>% 
  select(method, everything())
cv_yearly
write.csv(cv_yearly,  "03-data/code04_timeseries_cv_yearly.csv")
```



```{r}
emean_total=list()
emean_total[[1]]=mean_ets[[1]]
for (x in 1:1119) {
  emean_total[[x+1]]= c(emean_total[[x]], mean_ets[[x+1]])
}


```


mean error

```{r}

mae=mean(emean_total[[1120]], na.rm=TRUE)
mae
```


# rmse error

```{r}
for (x in 1:1120) {
  e_ets[[x]]=   e_ets[[x]]^2
}
e_ets[[1]]
```


```{r}
e_total=list()
e_total[[1]]=e_ets[[1]]
for (x in 1:1119) {
  e_total[[x+1]]= c(e_total[[x]], e_ets[[x+1]])
}
#e_total[[1120]]
```

```{r}
rmse= mean(e_total[[1120]], na.rm=TRUE)
rmse= sqrt(rmse)


```

the rmse is sqrt(11.12473)


# saving rmse and mae in a data frame

```{r}
cv <-data.frame(method=c("ETS"), MAE= mae, RMSE =rmse)
cv
```


```{r}
write.csv(cv,  "03-data/code04_timeseries_cv.csv")
```



END