---
title: "arima forecast"
output: 
  html_notebook:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
    number_sections: true
    code_folding: "hide"
    theme: "cosmo"
    highlight: "monochrome"
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
#install.packages("devtools")
library(devtools)
library(tidyverse)
library(summarytools)
library(xtable)
library(knitr)
library(ExPanDaR)
library(plotly)
library(REAT)
library(ggpubr)
library(DT)
library(ggrepel)
library(forecast)

# Change the presentation of decimal numbers to 4 and avoid scientific notation
library(readxl)
options(prompt="R> ", digits=7, scipen=999)
```

# loading data

```{r}
library(readr)
long_master_nona <- read_csv("01-raw-data/noNA_long_all_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"),    col_types = cols(NDVr = col_double(), 
       eb_ndvr = col_double(), eb_npir = col_double(), 
       nimr = col_double(), Litr= col_double(), 
        nmrpr = col_double(), nmpr = col_double() ))

wide_master_nona <- read_csv("01-raw-data/noNA_wide_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"))


cv<- read_csv(  "03-data/code04_timeseries_cv.csv")
cv<-cv[,-1]

cvy<- read_csv(  "03-data/code04_timeseries_cv_yearly.csv")
cvy<-cvy[,-1]
```

```{r}
cvy
```

# creating an ID 1:1120 for each municipality

```{r}
nmr<- long_master_nona %>% 
  filter(year>=2003) %>% 
  select(2:3,year, starts_with("eb_nmr")) %>% 
  arrange(code) %>% 
  mutate(code=as.character(code))
#nmr %>% 
a<- nmr %>% 
  filter(year==2010) %>% 
  select(code) %>% 
  mutate(ID=1:nrow(.))
a
nmr<- left_join(nmr,a,by="code")
nmr
```

# creaing a list of time series

```{r}
ts_list=list()
id=c(1:1120)

for (x in seq_along(id)) {
  try <- nmr %>% 
  filter(ID==id[x])
  
  ts_list[[x]] <- ts(try$eb_nmr, start=2003)
}

```

# plotting the forecast for 2020 for one municipality

```{r}

id=1

autoplot(ts_list[[id]])

 auto.arima(ts_list[[id]]) %>% 
forecast(h=4) %>% 
    autoplot()
  
ggAcf(ts_list[[id]])

Box.test(ts_list[[id]], lag = 5, type = "Ljung")

```

    
# Estimating an arima model for the series of each municipality 
and predicting crime for 2022
    
    
```{r}
nmr2022=c()

for (x in 1:1120) {
  fit <- auto.arima(ts_list[[x]])
#summary(fit)
 foreca <-fit %>% forecast(h=4)
nmr2022[x]= data.frame(foreca)[4,1]
 }

```
    
    
According to this rough prediction how many municipalities will reach the 2022 target?
    
```{r}
nmr2018<- nmr %>% filter(year==2018)

nmr2018$pred2022 <- nmr2022
nmr2018

target<- 10000-(23.23/10)
target


nmr2018 %>% 
  filter(eb_nmr<target)

nmr2018 %>% 
  filter(pred2022<target)

nmr2018 %>% arrange(desc(pred2022))

```


# comparing arima and exponential smoothing models


```{r}

# Set up forecast functions for ETS and ARIMA models
fets <- function(x, h) {
  forecast(ets(x), h = h)
}
farima <- function(x, h) {
  forecast(auto.arima(x), h=h )
}
```

# runnning this may take a few minutes

```{r}

MSE_ets=c()
MSE_arima=c()
for (xx in 1:1120) {
# Compute CV errors for ETS on each time series as e1
e1 <- tsCV(ts_list[[xx]], fets, h=1)

# Compute CV errors for ARIMA on each time series as e2
e2 <- tsCV(ts_list[[xx]], farima,h=1)

# Find MSE of each model class
MSE_ets[xx] <-  mean(e1^2, na.rm=TRUE)
MSE_arima[xx] <- mean(e2^2,  na.rm=TRUE)
}
```


It appears that arima is the model that has the best forecasting power for most municipalities


```{r}
length(MSE_arima)
length(MSE_ets)
MSE<-data.frame(MSE_arima, MSE_ets) %>% 
  mutate(arima_best= ifelse(MSE_arima<MSE_ets,1,0 )) %>% 
  mutate(equal=ifelse(round(MSE_arima,4)==round( MSE_ets,4),1,0))

         
MSE %>%          
group_by(arima_best) %>% 
  summarise(n=n())

MSE %>%          
group_by(equal) %>% 
  summarise(n=n())

MSE

```


# creating an arima prediction for 2018 with data up to 2014

```{r}
fc_arima2018=c()

for (x in 1:1120) {
  fit <- auto.arima(window(ts_list[[x]], end=2014))
#summary(fit)
 foreca <-fit %>% forecast(h=4)
fc_arima2018[x]= data.frame(foreca)[4,1]
 }
```



```{r}
nmr_2018_arima <- nmr %>% 
  filter(year==2018)

nmr_2018_arima$fc_arima2018 <- fc_arima2018

nmr_2018_arima <- nmr_2018_arima  %>%
  mutate(forecast_error=eb_nmr - fc_arima2018)
```


```{r}
nmr_2018_arima %>%  arrange(forecast_error)
mean((nmr_2018_arima$forecast_error)^2)
```



#  time series cross validation (h=4)

Running this chunk may take a few minutes.

```{r}
e_arima=list()
for (jj in 1:1120) {
  e<- tsCV(ts_list[[jj]], initial = 4,  farima,h=4)
e_arima[[jj]]<- e[,4]
}

```



# new table


```{r}
new_arima=list()
for (i in 1:1120) {
  new_arima[[i]]=abs(e_arima[[i]])
}

for (xx in 1:1119) {
  new_arima[[xx+1]]= new_arima[[xx]]+new_arima[[xx+1]]
}
new_arima[[1120]]/1120
mean_arima<- (new_arima[[1120]]/1120)[5:12]
mean_arima
```

```{r}
new_arima=list()
for (i in 1:1120) {
  new_arima[[i]]=(e_arima[[i]])^2
}
for (xx in 1:1119) {
  new_arima[[xx+1]]= new_arima[[xx]]+new_arima[[xx+1]]
}
methodv=c("ARIMA (MAE)","ARIMA (RMSE)")
names=seq(2011,2018,1)
names= as.character(names)
new_arima[[1120]]/1120
rmse_arima<- (new_arima[[1120]]/1120)[5:12]
rmse_arima<- sqrt(rmse_arima)
rmse_arima
cv_yearly<- as.data.frame( rbind(mean_arima, rmse_arima))
colnames(cv_yearly)<-names
cv_yearly<- cv_yearly %>% 
  mutate(method=methodv) %>% 
  select(method, everything())
cv_yearly







cvy2<-rbind(cvy, cv_yearly)
cvy2
cv
write.csv(cvy2,  "03-data/code04_timeseries_cv_yearly.csv")
```



# mean error MAE

```{r}
e_arima[[1]]
mean_arima=list()
for (i in 1:1120) {
  mean_arima[[i]]=abs(e_arima[[i]])
}

emean_total=list()
emean_total[[1]]=mean_arima[[1]]
for (x in 1:1119) {
  emean_total[[x+1]]= c(emean_total[[x]], mean_arima[[x+1]])
}


```

```{r}

mae=mean(emean_total[[1120]], na.rm=TRUE)
mae
```


rmse error

```{r}
for (x in 1:1120) {
  e_arima[[x]]=   e_arima[[x]]^2
}

```


```{r}
e_total=list()
e_total[[1]]=e_arima[[1]]
for (x in 1:1119) {
  e_total[[x+1]]= c(e_total[[x]], e_arima[[x+1]])
}

```

```{r}
rmse=mean(e_total[[1120]], na.rm=TRUE)
rmse=sqrt(rmse)
```

the rmse is sqrt(8.55)

```{r}
cv<-rbind(cv, data.frame(method= c("ARIMA"), MAE= mae, RMSE=rmse))
cv
```

```{r}
write.csv(cv,  "03-data/code04_timeseries_cv.csv")
```



END