---
title: "classical convergence forecasting for nmr"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  html_notebook:
    code_folding: hide
    highlight: monochrome
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '4'
---


```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(devtools)
library(ExPanDaR)
library(knitr)
library(tidyverse)
library(modelr)
library(broom)
library(data.table)
library(readxl)

#Install and load older versions of the following packages
#remove.packages("REAT")
#install_version("REAT", version = "2.1.1", repos = "http://cran.us.r-project.org")
library(REAT)
library(forecast)

# Change the presentation of decimal numbers to 4 and avoid scientific notation
options(prompt="R> ", digits=8, scipen=999)
```

# Import the master dataste

Via the `tidyverse` package

```{r}
library(readr)
long_master_nona <- read_csv("01-raw-data/noNA_long_all_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"),    col_types = cols(NDVr = col_double(), 
       eb_ndvr = col_double(), eb_npir = col_double(), 
       nimr = col_double(), Litr= col_double(), 
        nmrpr = col_double(), nmpr = col_double() ))
wide_master_nona <- read_csv("01-raw-data/noNA_wide_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"))


cv<- read_csv(  "03-data/code04_timeseries_cv.csv")
cv<-cv[,-1]

cvy<- read_csv(  "03-data/code04_timeseries_cv_yearly.csv")
cvy<-cvy[,-1]
```

# Preparing the dataset

Selecting variables needed
```{r}
wide_master_nona
```


# how many municiplaities will achieve the SDGs target by 2022

is the speed of convergence changing over time?

```{r}

nmr<- wide_master_nona %>% 
  select(2:5,starts_with("eb_nmr"))
nmr
alpha=c()
lambda=c()
beta_coefficient=c()
half_life=c()
p_value=c()
years=c()
list_beta=list()
year_final=c()
for (i in 1:12) {
  tab<-nmr%>% 
  select(5:20)
  tab<- tab %>% 
    select(i,i+4)
  name1<- as.character(colnames(tab)[1])
  name2<- as.character(colnames(tab)[2])
  colnames(tab)[1]<-"FY_yo"
  colnames(tab)[2]<-"FY_yt"
    tab
beta <- betaconv.ols (tab$FY_yo, i, tab$FY_yt, i+4,
                       beta.plot = FALSE,
                       beta.plotLine = TRUE,
                       beta.plotX = paste("Ln (non murder rate in", name1), 
                       beta.plotY = paste("Annual Growth Rate",name1,name2), 
                       beta.plotTitle = paste(name1,name2),
                      beta.plotLineCol = "purple",output.results = FALSE)
alpha[i]<-as.data.frame(beta[["abeta"]])[1,1]
beta_coefficient[i]<-as.data.frame(beta[["abeta"]])[2,1]
lambda[i]<-as.data.frame(beta[["abeta"]])[3,1]
half_life[i]<-as.data.frame(beta[["abeta"]])[4,1]
p_value[i]<-as.data.frame(beta[["abeta"]])[2,4]
years[i]<- paste(substr(name1,nchar(name1)-3,nchar(name1)),substr(name2,nchar(name2)-3,nchar(name2)),sep="-")
year_final[i]<- as.integer(substr(name2,nchar(name2)-3,nchar(name2)),sep="-")
list_beta[[i]]<- beta
}

nmr_beta<- data.frame(years,year_final, alpha,beta_coefficient,lambda, half_life,p_value)
```

```{r}
nmr_beta
```

```{r}
nmr_beta
nmr_beta %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=alpha)) +
  geom_point()+
  geom_smooth(method="lm")
nmr_beta %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")

nmr_beta %>% 
  ggplot(aes(x=year_final, y=beta_coefficient))+
  #ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")

lm_nmr_beta <- lm(beta_coefficient~year_final, nmr_beta)
lm_nmr_beta %>%  tidy()
newd <- data.frame(year_final=2022)
beta_2022<- predict(lm_nmr_beta,newd)

lm_nmr_lambda <- lm(lambda~year_final, nmr_beta)
lm_nmr_lambda %>%  tidy()
newd <- data.frame(year_final=2022)
lambda_2022<- predict(lm_nmr_lambda,newd)

lm_nmr_alpha <- lm(alpha~year_final, nmr_beta)
lm_nmr_alpha %>%  tidy()
newd <- data.frame(year_final=2022)
alpha_2022<- predict(lm_nmr_alpha,newd)
alpha_2022
beta_2022
lambda_2022
```

```{r}
final_coeff<- nmr_beta %>%  select(-1)
final_coeff
```

# saving estimates of coefficients over time 

```{r}
write.csv(final_coeff, "03-data/code04_table_pred.csv")
```


# how many municiplaities will achieve the NMR SDGs target by 2022


```{r}
#long_master_nona
long_master_nona %>% 
  select(1:6,eb_nmr) %>% 
  filter(year==2018) %>% 
  summarise(mean(eb_nmr))

#target 23,23 per 100.000 people

target<- 10000-(23.23/10)
target

nmr_2018 <- long_master_nona %>% 
  select(code,year,eb_nmr) %>% 
  filter(year==2018) %>% 
  filter(eb_nmr < target)

nmr_2018

y2022=c()
y2022 <- exp(alpha_2022+((1+beta_2022)*log(nmr_2018$eb_nmr)))
#y2022

nmr_2018$NMr2022 <- y2022
nmr_2018 %>% 
  filter(NMr2022<target)
```


We would like to forecast the value of NMR for the year 2022 but having data up to 2018, that is h=4.
we could try to use as a training set data from 2003 to 2014 in order to predidct the value of 2018 and compare it with the real value


```{r}
nmr_beta_14<- nmr_beta %>% 
  filter(year_final<=2014)

nmr_beta_14
nmr_beta_14 %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=alpha)) +
  geom_point()+
  geom_smooth(method="lm")
nmr_beta_14 %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")

nmr_beta_14 %>% 
  ggplot(aes(x=year_final, y=beta_coefficient))+
  #ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")

lm_nmr_beta_14 <- lm(beta_coefficient~year_final, nmr_beta_14)
lm_nmr_beta_14 %>%  tidy()
newd <- data.frame(year_final=2018)
beta_2018<- predict(lm_nmr_beta_14,newd)

lm_nmr_lambda <- lm(lambda~year_final, nmr_beta_14)
lm_nmr_lambda %>%  tidy()
newd <- data.frame(year_final=2018)
lambda_2018<- predict(lm_nmr_lambda,newd)

lm_nmr_alpha <- lm(alpha~year_final, nmr_beta_14)
lm_nmr_alpha %>%  tidy()
newd <- data.frame(year_final=2018)
alpha_2018<- predict(lm_nmr_alpha,newd)
alpha_2018
beta_2018
lambda_2018
```


We can now compute the forecast for 2018 and compare it to the actual value


```{r}
#long_master_nona

nmr_2018 <- wide_master_nona %>% 
  select(code,eb_nmr_2014, eb_nmr_2018) 

nmr_2018

y2018=c()
y2018 <- exp(alpha_2018+((1+beta_2018)*log(nmr_2018$eb_nmr_2014)))
#y2022

nmr_2018$fc2018 <- y2018
nmr_2018 <- nmr_2018 %>% 
  mutate(forecast_error=eb_nmr_2018-fc2018)
nmr_2018 %>% arrange(forecast_error)

```


```{r}
mean((nmr_2018$forecast_error)^2)
```



#  time series cross validation for h=4
```{r}
nmr_beta
```

```{r}
alphax=c()
betax=c()
lambdax=c()
nmr_beta
for (x in seq_along(2011:2018)) {
  ye=2007+(x-1)
  nmr_betax<- nmr_beta %>% 
  filter(year_final<=ye)

lm_nmr_betax <- lm(beta_coefficient~year_final, nmr_betax)
lm_nmr_betax %>%  tidy()
newd <- data.frame(year_final=ye+4)
beta_x4<- predict(lm_nmr_betax,newd)

lm_nmr_lambda <- lm(lambda~year_final, nmr_betax)
lm_nmr_lambda %>%  tidy()
newd <- data.frame(year_final=ye+4)
lambda_x4<- predict(lm_nmr_lambda,newd)

lm_nmr_alpha <- lm(alpha~year_final, nmr_betax)
lm_nmr_alpha %>%  tidy()
newd <- data.frame(year_final=ye+4)
alpha_x4<- predict(lm_nmr_alpha,newd)
alphax[x]= alpha_x4
betax[x]=beta_x4
lambda[x]=lambda_x4
 
}


```

```{r}
alphax
```


```{r}
nmr_xx <- wide_master_nona %>% 
  select(code,starts_with("eb_nmr")) 

nmr_xx

fc_nmr=c()

for (jj in 1:8) {
  
  fc_nmr <- exp(alphax[jj]+((1+betax[[jj]])*log(nmr_xx[[jj+5]])))
  
  nmr_xx<- cbind(nmr_xx, fc_nmr)
  colnames(nmr_xx)[ncol(nmr_xx)]<-paste("fc",colnames(nmr_xx)[jj+9],sep="")
  
}

nmr_xx
```


```{r}
nmr_xxx<- nmr_xx %>% 
  select(10:25)
year=c(2011:2018)
forecast_error=c()
nmr_xxx
for (i in seq_along(year)) {
   forecast_error= (nmr_xxx[[i]]-nmr_xxx[[i+8]])^2
nmr_xxx<- cbind(nmr_xxx, forecast_error)
colnames(nmr_xxx)[ncol(nmr_xxx)]<- paste("for.error", as.character(year[i]), sep="")
   }

```

```{r}
nmr_xxx
error<-nmr_xxx %>% 
  select(17:24)
rmsev<-mean(as.matrix(error))
rmsev= sqrt(rmsev)
```

# new tables

```{r}
#cv
#RMSE
#error
rmse<-error %>% 
  summarise_all(sum)
rmse<-sqrt(rmse/1120)
rmse

#MAE

mae_y <- sqrt(error)
#mae_y
mae_y<- mae_y%>% 
  summarise_all(sum)
mae_y<- mae_y/1120
mae_y

cvy
cv_yearly<- rbind(mae_y, rmse)

methodv=c("BETA (MAE)","BETA (RMSE)")
names=seq(2011,2018,1)
names= as.character(names)
colnames(cv_yearly)<-names
cv_yearly<- cv_yearly %>% 
  mutate(method=methodv) %>% 
  select(method, everything())
cv_yearly
cvy2<- rbind(cvy, cv_yearly)
write.csv(cvy2,  "03-data/code04_timeseries_cv_yearly.csv")
```





the RMSE is  sqrt(6.1883245)

```{r}
error
mae<-mean(as.matrix(sqrt(error)))
mae
```

```{r}
cv<-rbind(cv, data.frame(method= c("Beta"), MAE= mae, RMSE=rmsev))
cv
```

```{r}
write.csv(cv,  "03-data/code04_timeseries_cv.csv")
```


# Improving the prediction of alpha and beta 
```{r}
nmr_beta 
nmr_beta %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=alpha)) +
  geom_point()+
  geom_smooth(method="lm")
nmr_beta %>% 
  #ggplot(aes(x=year_final, y=beta_coefficient))+
  ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")

nmr_beta %>% 
  ggplot(aes(x=year_final, y=beta_coefficient))+
  #ggplot(aes(x=year_final, y=lambda)) +
  geom_point()+
  geom_smooth(method="lm")
```


for forecasting nmr a prediction of beta and apha is needed can the forecasting of these coefficients be improved?


```{r}
alpha_ts<- ts(nmr_beta$alpha, start=2007)
beta_ts<- ts(nmr_beta$beta_coefficient, start=2007)
autoplot(alpha_ts)
autoplot(beta_ts)

ggAcf(alpha_ts)
ggAcf(beta_ts)

fita_arima <- auto.arima(alpha_ts)
fitb_arima <- auto.arima(beta_ts)
fita_ets <- ets(alpha_ts)
fitb_ets <- ets(beta_ts)


```

as it can be seen below, beta and alpha may not be predicted using arima or ets models since the trend is not considered. 


```{r}
fita_arima %>% forecast(h=4) %>% autoplot()
fitb_arima %>% forecast(h=4) %>% autoplot()
fita_ets %>% forecast(h=4) %>% autoplot()
fitb_ets %>% forecast(h=4) %>% autoplot()


```


```{r}
sessionInfo()
```