---
title: "03-d-spatial-forecasting"
author: "Felipe Santos-Marquez"
date: "4/12/2020"
output: html_document
---

In this file we look for the tendencies of the beta, alpha and theta coeffcients of the beta regressions (spatial lag of x model) 

# loading required packages

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(RCurl)
library(tidyverse) #
library(sp)
library(sf) #
library(fastmap)
library(tmap)
library(tmaptools)
library(spdep) #
library(rgeos)
library(readr)
library(rgdal)
library(broom)
library(stargazer)
library(skimr)
library(spatialreg) #
library(forecast)
# Change the presentation of decimal numbers to 4 and avoid scientific notation
options(prompt="R> ", digits=5, scipen=999)
```

# Importing the master data

```{r}
noNA_all <- read_csv("01-raw-data/noNA_wide_data_master_health_violence_services_education.csv", locale = locale(encoding = "ISO-8859-1"))

noNA_all

cv<- read_csv(  "03-data/code04_timeseries_cv.csv")
cv<-cv[,-1]
cv

cvy<- read_csv(  "03-data/code04_timeseries_cv_yearly.csv")
cvy<-cvy[,-1]
```

# Import the municipalities shapefile


```{r}

mun<- st_read("01-raw-data/master/mun_master_thesis.shp")
mun
object.size(mun)
s.sp <- as(mun, "Spatial")
class(s.sp)
class(mun)

```


# selecting intial and final years and creating the variables for the regressions

```{r}
colnames(noNA_all)

master<- noNA_all %>% 
  select(c(2:4), starts_with("eb_nmr"))
master
```

taking the log of all variables in the master dataframe
since log(0) is not defined the it is replaced by 0.000001

```{r}
master[master == 0] <- 0.000001

a<- log(master[,-1:-3])
a
b<- master[,1:3] 
master_log<- cbind(b,a)

```

computing the groewth rate as the substraction Log(yt)-log(y0)

```{r}
master_log
seq=c()
seq<- seq(8,19)

for (i in seq) {
  master_log[,12+i]<-  master_log[,i]-master_log[,i-4]
  old_name<-  colnames(master_log)[i]
  colnames(master_log)[12+i]<- paste("gr",old_name,sep="_")
}
master_log
nmr18masterlog <- master_log %>%  select(code,eb_nmr_2018)
```

Showing the classical beta conevrgene for one variables

```{r}
master_log <- master_log%>% 
  select(-16:-19)
```

```{r}
master_log
```


```{r}
master_log %>% 
  ggplot(aes(x=eb_nmr_2003,y=gr_eb_nmr_2007))+
  geom_point()+
  geom_smooth(method = "lm",se=0)

master_log<- master_log %>% 
  select(1:3, 4,16, 5,17,6,18,7,19,8,20,9,21,10,22,11,23,12,24,13,25,14,26,15,27 )
```


merging data and shape-file

```{r}
mun_merge<- merge(s.sp, master_log, by.x="MPIO_CCNCT", by.y="code")
mun_merge2<- merge(s.sp, nmr18masterlog, by.x="MPIO_CCNCT", by.y="code")
```

```{r}
names(mun_merge2) 
```

# creating weight matrix- queen contiguity criteria

```{r}
#ids<- mun_merge@data$MPIO_CCNCT
#IDS <- as.character(ids)
#nb <- poly2nb(mun_merge, queen=TRUE, row.names = IDS)
#W.matrixpol <- nb2listw(nb, style="W", zero.policy=TRUE)
#summary(W.matrixpol)

```

Exporting weight matrix

```{r}
#queen<- listw2sn(W.matrixpol)
#summary(W.matrixpol)
#summary(queen)
#write.sn2gwt(listw2sn(W.matrixpol), "data/shapefile/master/queen-contiguity.gwt" )

#nc.5nn.mat <- nb2mat(nb)

#write.csv(nc.5nn.mat, file="data/shapefile/master/queen-contiguity.csv" )

```


```{r}
#W.matrixpol<- read.gwt2nb("data/shapefile/master/queen-contiguity.gwt")

W.matrixpol.mat <-read.geoda("01-raw-data/master/queen-contiguity.csv",row.names = 1 )

mat<- as.matrix(W.matrixpol.mat)
#View(mat)


W.matrixpol<- mat2listw(mat)

lw <- nb2listw(W.matrixpol$neighbours, style = "W")

moran.plot(mun_merge$eb_nmr_2003, listw = W.matrixpol)

moran.plot(mun_merge$eb_nmr_2003, listw = lw)

#class(lw)

#class(W.matrixpol)

#lw$neighbours
#W.matrixpol$neighbours
```

```{r}
mun_merge@data
#mun_merge@data[,13]

```


```{r}
 seq<- seq(14,36,by=2)
seq

reg.eq=list()
for (i in 14:36) {
  reg.eq[[i+1]] <-  colnames(mun_merge@data)[i-1]
}

reg.eq[[14]] <- gr_eb_nmr_2007 ~ eb_nmr_2003
reg.eq[[16]] <- gr_eb_nmr_2008 ~ eb_nmr_2004
reg.eq[[18]] <- gr_eb_nmr_2009 ~ eb_nmr_2005
reg.eq[[20]] <- gr_eb_nmr_2010 ~ eb_nmr_2006
reg.eq[[22]] <- gr_eb_nmr_2011 ~ eb_nmr_2007
reg.eq[[24]] <- gr_eb_nmr_2012 ~ eb_nmr_2008
reg.eq[[26]] <- gr_eb_nmr_2013 ~ eb_nmr_2009
reg.eq[[28]] <- gr_eb_nmr_2014 ~ eb_nmr_2010
reg.eq[[30]] <- gr_eb_nmr_2015 ~ eb_nmr_2011
reg.eq[[32]] <- gr_eb_nmr_2016 ~ eb_nmr_2012
reg.eq[[34]] <- gr_eb_nmr_2017 ~ eb_nmr_2013
reg.eq[[36]] <- gr_eb_nmr_2018 ~ eb_nmr_2014

reg.eq[[14]]
```

```{r}
reg=list()
for (i in seq) {
  reg[[i]]<-lm(reg.eq[[i]], data = mun_merge)
}
lmmorantest=list()
for (i in seq) {
 lmmorantest[[i]] <- lm.morantest(reg[[i]], lw)
}

lmmorantest[[14]]

```

 
Model SLX

Spatially Lagged X y=Xß+TWX+e , T=theta, and L=lambda

```{r}

sar=list()
for (i in seq) {
  sar[[i]] <- lmSLX(reg.eq[[i]],data = mun_merge, listw = lw)
}

```

# lw instead of W.matrixpol should be used

```{r}

#lw <- spdep::nb2listw(nb, style="W")
COL.SLX <- lmSLX(gr_eb_nmr_2007 ~ eb_nmr_2003, data= mun_merge, listw=lw)
summary(COL.SLX)


COL.SLX <- lmSLX(gr_eb_nmr_2007 ~ eb_nmr_2003, data= mun_merge, listw=W.matrixpol)
summary(COL.SLX)
```

```{r}
summary(sar[[14]])
```

choose variable for which to diplay sem


```{r}
names.dep=names(mun_merge[,-ncol(mun_merge)])
i=14
substr(names.dep[i-1],1,nchar(names.dep[i-1])-4)

a<-summary(sar[[14]])
#sar[[14]]$coefficients[3]
a<-as.data.frame(a$coefficients)
a
a[3,4]
summary(sar[[14]])
names(sar[[14]]$coefficients)[3]
```

 # changing names of coefficients and formula
 
```{r}
for (i in seq) {
  names(sar[[i]]$coefficients)[3]<-"theta"
  names(sar[[i]]$coefficients)[2]<-"log Y0"
  names(reg[[i]]$coefficients)[2]<- "log Y0"
  names(sar[[i]]$coefficients)[1]<-"alpha"
  names(reg[[i]]$coefficients)[1]<-"alpha"
  }



 for (i in seq) {

sar[[i]]$call$formula[2]<-substr(names.dep[i-1],1,nchar(names.dep[i-1])-4)
reg[[i]]$call$formula[2]<-substr(names.dep[i-1],1,nchar(names.dep[i-1])-4)
 }


```
 
 
```{r}
reg[[14]]
sar[[14]]
```
 
 

```{r}
 vect =rep(1:3, 6)
columns = rep(c("\\shortstack{no spatial \\\\ effects}", "\\shortstack{spatial\\\\  lag}"),6)

column.labels  = c("no spatial \n effects", "spatial  \nlag")
columns
```



```{r}
reg[[14]]$coefficients[2]
sar[[14]]$coefficients[2]
```

```{r}
years<-substr(names.dep, nchar(names.dep)-3, nchar(names.dep))
years<- as.integer(years)
years
```

```{r}
names.dep=names(mun_merge)
names.dep
years<-substr(names.dep, nchar(names.dep)-3, nchar(names.dep))
years<- as.integer(years)
years
```


```{r}
seq

speed.values=c()
halflife.values=c()
speed.values.sar=c()
halflife.values.sar=c()


for (x in seq) {
  beta <- reg[[x]]$coefficients[2]
speed.values[x]<- -log(1+beta)/(years[x]-years[x-1])
halflife.values[x]<- log(2)/  speed.values[x]


beta <- sar[[x]]$coefficients[2]
speed.values.sar[x]<- -log(1+beta)/(years[x]-years[x-1])
halflife.values.sar[x]<- log(2)/  speed.values.sar[x]


}


data.frame(speed.values, halflife.values, names.dep) %>% 
  filter(!is.na(halflife.values)) %>% 
  arrange(names.dep)

data.frame(speed.values, halflife.values, names.dep)[seq,]


#speed.values[seq]
#speed.values.sem[seq]
#speed.values.sar[seq]
all.speed.values<- c(rbind(speed.values[seq],speed.values.sar[seq]))

all.halflife.values<- c(rbind(halflife.values[seq], halflife.values.sar[seq]))
```

```{r}
all.speed.values
all.halflife.values

```



# model selection lines



#lines to be added to stargazer

```{r}
halflife=c("half life", round(all.halflife.values,2))
speed=c("\\shortstack{speed of\\\\ convergence}", round(all.speed.values,3))


columns
halflife
speed
```

# stargazer tables

```{r}
i=14
table.positions=c(1,2:13)

writeLines(capture.output(stargazer(reg[[i]], sar[[i]],reg[[i+2]], sar[[i+2]],reg[[i+4]], sar[[i+4]],reg[[i+6]], sar[[i+6]],reg[[i+8]], sar[[i+8]],reg[[i+10]], sar[[i+10]],digits = 2, float.env = "sidewaystable", column.sep.width = "0.1pt",dep.var.labels.include = TRUE, model.numbers = TRUE, model.names = FALSE, column.labels= columns, column.separate = rep(1,13), add.lines = list(speed[table.positions], halflife[table.positions]), covariate.labels = c("$\\alpha$", "$Y_{T0}$"), omit.stat = c( "ser","sigma2", "wald", "f"))), "03-data/table_beta_nmr_pred1.tex")

i=26
table.positions=c(1,14:25)

writeLines(capture.output(stargazer(reg[[i]], sar[[i]],reg[[i+2]], sar[[i+2]],reg[[i+4]], sar[[i+4]],reg[[i+6]], sar[[i+6]],reg[[i+8]], sar[[i+8]],reg[[i+10]], sar[[i+10]],digits = 2, float.env = "sidewaystable", column.sep.width = "0.1pt",dep.var.labels.include = TRUE, model.numbers = TRUE, model.names = FALSE, column.labels= columns, column.separate = rep(1,13), add.lines = list(speed[table.positions], halflife[table.positions]), covariate.labels = c("$\\alpha$", "$Y_{T0}$"), omit.stat = c( "ll","lr", "ser","sigma2", "wald", "f"))), "03-data/table_beta_nmr_pred2.tex")

```



```{r}
theta=c()
for (x in seq_along(seq)) {
  theta[x]<- sar[[seq[x]]]$coefficients[3]
}

theta

final_coeff <- data.frame(theta, finalyear=2007:2018)

final_coeff%>% 
  ggplot(aes(x=finalyear, y=theta))+
  geom_point()+
  geom_smooth(method="lm")


final_coeff

```

```{r}
beta_nmr=c()
alpha_nmr=c()
for (i in 1:12) {
 beta_nmr[i]<-  sar[[seq[i]]]$coefficients[2]
 alpha_nmr[i]<- sar[[seq[i]]]$coefficients[1]
}

final_coeff<- data.frame(final_coeff, alpha_nmr, beta_nmr) 
final_coeff
final_coeff_wide<- final_coeff
```

```{r}
final_coeff%>% 
  #filter(finalyear<=2017) %>% 
  ggplot(aes(x=finalyear, y=theta))+
  geom_point()+
  geom_smooth(method="lm")

final_coeff%>% 
  ggplot(aes(x=finalyear, y=alpha_nmr))+
  geom_point()+
  geom_smooth(method="lm")

final_coeff%>% 
  ggplot(aes(x=finalyear, y=beta_nmr))+
  geom_point()+
  geom_smooth(method="lm")


beta_pred=0
alpha_pred=0
theta_pred=0


reg_beta<-lm(beta_nmr~finalyear, final_coeff)
#summary(reg)
beta_pred<- predict(reg_beta, data.frame(finalyear=2022))
reg_alpha<-lm(alpha_nmr~finalyear, final_coeff)
#summary(reg2)
alpha_pred<- predict(reg_alpha, data.frame(finalyear=2022))
reg_theta<-lm(theta~finalyear, final_coeff)
#summary(reg)
theta_pred<- predict(reg_theta, data.frame(finalyear=2022))

beta_pred
alpha_pred
theta_pred
```


 
```{r}
final_coeff
colnames(final_coeff)[3]<-"alpha"
colnames(final_coeff)[4]<-"beta"


final_coeff<- final_coeff %>% 
  select(2,1,3,4) %>% 
  gather(beta_con, value,2:4) 

final_coeff%>% 
  ggplot(aes(x=finalyear, y=value)) +
  geom_line(linetype = "dashed", lwd=1)+
  geom_smooth(method="lm", se=FALSE)+
  facet_grid(rows=vars(beta_con),scales="free")+
  theme_minimal() +
  labs(subtitle = "",
       x = "final year",
       y = "value of the coefficient") +
  theme(text=element_text( family="Palatino"))
```
 
```{r}
write.csv(final_coeff, "03-data/code04_table_spatialpred.csv")
```


# 2022 target

```{r}

tarnmr<- 10000-(23.23/10)
tarnmr
#exp(predict(sar[[36]], as.data.frame(mun_merge@data$NMr2014), listw = W.matrixpol)+ log(mun_merge@data$NMr2014))
```
     1 
-0.5049 
       1 
-0.53483 
      1 
0.56297 

# 2022 crime forecast


```{r}
#mun_merge@data$MPIO_CCNCT
try <- data.frame(code = mun_merge2@data$MPIO_CCNCT, nmr18 = mun_merge2@data$eb_nmr_2018)

try$nmrlag18<-lag.listw(lw,
                         mun_merge2@data$eb_nmr_2018, zero.policy=TRUE)
try


x=1
try_alpha<- sar[[seq[x]]]$coefficients[1]
try_beta<-sar[[seq[x]]]$coefficients[2]
try_theta<- sar[[seq[x]]]$coefficients[3] 
try_alpha
try_beta
try_theta

try_alpha <- alpha_pred
try_beta <- beta_pred
try_theta <- theta_pred

try_alpha
try_beta
try_theta
try$nmr18nl<-  exp(try$nmr18)
try$NMr2022<-exp(try_alpha+((1+try_beta)*(try$nmr18))+(try_theta*(try$nmrlag18)))

try

```



```{r}

sum(try$nmr18nl< tarnmr)
sum(try$NMr2022< tarnmr)

pred2022<- master %>% 
  select(code, eb_nmr_2003, eb_nmr_2018) %>% 
  left_join(., try, by="code") %>% 
  select(code,eb_nmr_2003, eb_nmr_2018, NMr2022 )
pred2022
write.csv(pred2022, "03-data/code04_table_pred2022.csv")

data.frame(descr(pred2022,
      headings = FALSE, # remove headings
      stats = "common"))# most common descriptive statistics
```




# if we have data up to 2014 what will the forecast for nmr in 2018 be?

```{r}
final_coeff_wide
final_coeff_14 <- final_coeff_wide %>% filter(finalyear<=2014)
final_coeff_14


final_coeff_14%>% 
  ggplot(aes(x=finalyear, y=alpha_nmr))+
  geom_point()+
  geom_smooth(method="lm")

final_coeff_14%>% 
  ggplot(aes(x=finalyear, y=beta_nmr))+
  geom_point()+
  geom_smooth(method="lm")

final_coeff_14%>% 
  ggplot(aes(x=finalyear, y=theta))+
  geom_point()+
  geom_smooth(method="lm")

beta_pred=0
alpha_pred=0
rho_pred=0


reg_beta<-lm(beta_nmr~finalyear, final_coeff_14)
#summary(reg)
beta_pred<- predict(reg_beta, data.frame(finalyear=2018))
reg_alpha<-lm(alpha_nmr~finalyear, final_coeff_14)
#summary(reg2)
alpha_pred<- predict(reg_alpha, data.frame(finalyear=2018))
reg_theta<-lm(theta~finalyear, final_coeff_14)
#summary(reg)
theta_pred<- predict(reg_theta, data.frame(finalyear=2018))


alpha_pred
beta_pred
theta_pred
```


forecast for 2018

```{r}
try <- data.frame(code = mun_merge@data$MPIO_CCNCT, eb_nmr_2014=mun_merge@data$eb_nmr_2014)
try$nmrlag14<-lag.listw(lw, mun_merge@data$eb_nmr_2014,zero.policy=TRUE)
try<- left_join(try, nmr18masterlog, by="code")

try
```


```{r}
try
try_alpha <- alpha_pred
try_beta <- beta_pred
try_theta <- theta_pred
try<- try %>% 
  mutate(fc2018=exp(try_alpha+((1+try_beta)*(try$eb_nmr_2014))+(try_theta*(try$nmrlag14)))) %>% 
  mutate(eb_nmr2018=exp(eb_nmr_2018)) %>% 
    mutate(forecast_error=eb_nmr2018 - fc2018)
try
```

```{r}
try %>% arrange(forecast_error)
mean((try$forecast_error)^2)
```

# time series cross validation for h=4


```{r}

 final_coeff <- final_coeff_wide
betax=c()
alphax=c()
thetax=c()
for (x in seq_along(2011:2018)) {
  ye=2007+(x-1)

final_coeff_xx <- final_coeff %>% filter(finalyear<=ye)

reg_beta<-lm(beta_nmr~finalyear, final_coeff_xx)
beta_pred<- predict(reg_beta, data.frame(finalyear=ye+4))

reg_alpha<-lm(alpha_nmr~finalyear, final_coeff_xx)
alpha_pred<- predict(reg_alpha, data.frame(finalyear=ye+4))

reg_theta<-lm(theta~finalyear, final_coeff_xx)
theta_pred<- predict(reg_theta, data.frame(finalyear=ye+4))

betax[x] = beta_pred
alphax[x]= alpha_pred
thetax[x]= theta_pred
}
```


```{r}
thetax
alphax
betax
master
mun_merge_nolog<- merge(s.sp, master, by.x="MPIO_CCNCT", by.y="code")
```

```{r}
masterxx<- master
mun_merge_nolog@data
yearx=c(2007:2018)
for (xx in 17:28) {
  mun_merge_nolog@data$nmrlag <-lag.listw(lw, mun_merge_nolog@data[[xx]],zero.policy=TRUE)
  a<- mun_merge_nolog@data %>%  select(MPIO_CCNCT, nmrlag)
masterxx <- left_join(masterxx, a, by=c("code"="MPIO_CCNCT"))
colnames(masterxx)[ncol(masterxx)]<- paste("lag",colnames(mun_merge_nolog@data)[xx], sep = "")
}

```

```{r}
masterxx
masterxx<- masterxx %>% 
  select(8:31)
masterxx
```


```{r}
masterxx
j=1
for (jj in 1:8) {
  fc_nmr <-  exp(alphax[jj]+((1+betax[[jj]])*(log(masterxx[[jj]])))+(thetax[[jj]]*(log(masterxx[[jj+12]]))))
  masterxx<- cbind(masterxx, fc_nmr)
  colnames(masterxx)[ncol(masterxx)]<-paste("fc",colnames(masterxx)[jj+4],sep="")
  
}
```

```{r}
 masterxx
```


```{r}
#log(masterxx[[jj]])

masterxxx<- masterxx %>% 
  select(starts_with(c("eb","fc"))) %>% 
  select(5:20)
masterxxx

year=c(2011:2018)
forecast_error=c()
for (i in seq_along(year)) {
   forecast_error= (masterxxx[[i]]-masterxxx[[i+8]])^2
masterxxx<- cbind(masterxxx, forecast_error)
colnames(masterxxx)[ncol(masterxxx)]<- paste("for.error", as.character(year[i]), sep="")
   }
```

```{r}
masterxxx

error<-masterxxx %>% 
select(starts_with("for.er"))
rmse = mean(as.matrix(error))
rmse = sqrt(rmse)
rmsev= rmse
error

```


# new tables


```{r}
#cv
#RMSE
#error
rmse<-error %>% 
  summarise_all(sum)
rmse<-sqrt(rmse/1120)
rmse
#MAE

mae_y <- sqrt(error)
#mae_y
mae_y<- mae_y%>% 
  summarise_all(sum)
mae_y<- mae_y/1120
mae_y

cvy
cv_yearly<- rbind(mae_y, rmse)

methodv=c("Spatial BETA (MAE)","Spatial BETA (RMSE)")
names=seq(2011,2018,1)
names= as.character(names)
colnames(cv_yearly)<-names
cv_yearly<- cv_yearly %>% 
  mutate(method=methodv) %>% 
  select(method, everything())
cv_yearly
cvy2<- rbind(cvy, cv_yearly)
cvy2
write.csv(cvy2,  "03-data/code04_timeseries_cv_yearly.csv")
```

the RMSE is  sqrt(5.5358)


```{r}
error
mae<-mean(as.matrix(sqrt(error)))
mae
```

```{r}
cv<-rbind(cv, data.frame(method= c("Beta Spatial"), MAE= mae, RMSE=rmsev))
cv
```

```{r}
write.csv(cv,  "03-data/code04_timeseries_cv.csv")
```

```{r}
masterxxx %>% 
  arrange(desc(for.error2018))
```

#can the predictions of alpha beta and theta be improved?


```{r}
final_coeff


alpha_ts<- ts(final_coeff$alpha_nmr, start=2007)
beta_ts<- ts(final_coeff$beta_nmr, start=2007)
theta_ts<- ts(final_coeff$theta, start=2007)

#alpha_ts<- window(alpha_ts,end=2017)
#beta_ts<- window(beta_ts,end=2017)
#rho_ts<- window(rho_ts,end=2017)

autoplot(alpha_ts)
autoplot(beta_ts)
autoplot(theta_ts)

ggAcf(alpha_ts)
ggAcf(beta_ts)
ggAcf(theta_ts)

fita_arima <- auto.arima(alpha_ts)
fitb_arima <- auto.arima(beta_ts)
fitr_arima <- auto.arima(theta_ts)
fita_ets <- ets(alpha_ts)
fitb_ets <- ets(beta_ts)
fitr_ets <- ets(theta_ts)

```

It appears that prediction models can not be used for these parameters, as the time series appear to be white noise. and the forecast resembles a naive forfecast

```{r}
fita_arima %>% forecast(h=4) %>% autoplot()
fitb_arima %>% forecast(h=4) %>% autoplot()
fitr_arima %>% forecast(h=4) %>% autoplot()

fita_ets %>% forecast(h=4) %>% autoplot()
fitb_ets %>% forecast(h=4) %>% autoplot()
fitr_ets %>% forecast(h=4) %>% autoplot()


```

